Revisiting the hydrodynamic wave modulation from linear wave theory

### 2024-08-23 paper/paper.tex
Comparing effective gravities

@@

Long waves induce vertical orbital accelerations, so the short
waves that ride on their surface move in an accelerated reference frame and
experience effective gravitational acceleration that is lower at the crests and
higher in the troughs \cite{longuet1986eulerian,longuet1987propagation}.
A long wave modulates gravitational acceleration at its surface:

\begin{equation}
\label{eq:gravity_modulation_general}
\widetilde{g}(\psi)
  = g \cos{\alpha} 
  + \dfrac{\partial W}{\partial t} \cos{\alpha}
  + \frac{\partial U}{\partial t} \sin{\alpha}
\end{equation}
where $\alpha = \arctan{\frac{\partial \eta}{\partial x}}$ is the local slope
of the long-wave surface.
Curvilinear effects on effective gravity can be removed by setting $\alpha = 0$.

For a linear wave, the effective gravity is:

@@

@@

\begin{equation}
\label{eq:eta_stokes}
\eta_{St} = a_L \left[
  \cos{\psi} +
  \dfrac{1}{2} \varepsilon_L \cos{2\psi} +
  \varepsilon_L^2 \left( \dfrac{3}{8} \cos{3\psi} - \dfrac{1}{16} \cos{\psi} \right)
\right]
\end{equation}
The corresponding gravitational acceleration at the surface of the Stokes wave
is, to the 4$^{th}$ order:

\begin{equation}
\label{eq:gravity_modulation_stokes}
\widetilde{g}(\psi) =
g \left\{
  1 - \varepsilon_L e^{k \eta_{St}}
  \left[ \cos{\psi} -
    \varepsilon_L \sin{\psi} \left(
      \sin{\psi}
      + \varepsilon_L \sin{2\psi}
      - \dfrac{1}{16} \varepsilon_L^2 \sin{\psi}
      + \dfrac{9}{8} \varepsilon_L^2 \sin{3\psi}
    \right)
  \right]
\right\}
\end{equation}

With the new long-wave modulated expressions for the short-wave wavenumber
(Eq. \ref{eq:k_short_exact}) and gravity
(Eqs. \ref{eq:gravity_modulation_linear}-\ref{eq:gravity_modulation_stokes}),
we can evaluate the long-wave modulated frequency as
$\widetilde{\omega}(\psi) = \sqrt{\widetilde{g} \widetilde{k}}$.
We compare the modulated intrinsic frequency for short waves on the surface of
linear and Stokes waves with $\varepsilon_L=2$ as function of phase, and their
phase-integral statistics in Fig. \ref{fig:frequency_modulation}.

>>>>

We now describe the analytical solutions for the short-wave wavenumber,
effective gravitational acceleration, amplitude, steepness, intrinsic frequency,
and intrinsic phase speed in presence of long waves.

@@

Long waves induce orbital accelerations, so the short waves that ride on their
surface move in an accelerated reference frame and experience effective
gravitational acceleration that is lower at the crests and higher in the troughs
\cite{longuet1986eulerian,longuet1987propagation}.
Further, a smaller but non-negligible effect is that of the centripetal force
due to the curvature of the long wave
\cite{phillips1981dispersion,zhang1990evolution}.
These effects are completely described by projecting the gravitational
acceleration from the coordinate that is perpendicular to the mean water surface
($z=0$) to that which is perpendicular to the long-wave surface ($z=\eta$),
and likewise for the long-wave orbital accelerations:

\begin{equation}
\label{eq:gravity_modulation_general}
\widetilde{g}
  = g \cos{\alpha} 
  + \dfrac{\partial W_{z=\eta}}{\partial t} \cos{\alpha}
  + \dfrac{\partial U_{z=\eta}}{\partial t} \sin{\alpha}
\end{equation}

\begin{equation}
\label{eq:local_slope}
\alpha = \tan^{-1}{\dfrac{\partial \eta}{\partial x}}
\end{equation}
where $\alpha$ is the local slope of the long-wave surface.
For a linear wave, the effective gravity is:

@@

Curvilinear effects on effective gravity can be neglected by setting $\alpha = 0$.

@@

\begin{equation}
\label{eq:eta_stokes}
\eta_{St} = a_L \left[
  \cos{\psi} +
  \dfrac{1}{2} \varepsilon_L \cos{2\psi} +
  \varepsilon_L^2 \left( \dfrac{3}{8} \cos{3\psi} - \dfrac{1}{16} \cos{\psi} \right)
\right]
\end{equation}
Without considering the curvilinear effects in Eq. \ref{eq:gravity_modulation_general},,
the gravitational acceleration at the surface of a Stokes wave is, to the
4$^{th}$ order:

\begin{equation}
\label{eq:gravity_modulation_stokes}
\widetilde{g}(\psi) =
g \left\{
  1 - \varepsilon_L e^{k \eta_{St}}
  \left[ \cos{\psi} -
    \varepsilon_L \sin{\psi} \left(
      \sin{\psi}
      + \varepsilon_L \sin{2\psi}
      - \dfrac{1}{16} \varepsilon_L^2 \sin{\psi}
      + \dfrac{9}{8} \varepsilon_L^2 \sin{3\psi}
    \right)
  \right]
\right\}
\end{equation}
The equivalent expression that includes the curvilinear effects can be derived
by evaluating orbital accelerations and local slope $\alpha$ at $z = \eta_{St}$,
and inserting them into Eq. \ref{eq:gravity_modulation_general}:

\begin{equation}
\label{eq:gravity_modulation_general_stokes}
\widetilde{g}
  = g \cos{\alpha} 
  + \dfrac{\partial W_{z=\eta_{St}}}{\partial t} \cos{\alpha_{St}}
  + \dfrac{\partial U_{z=\eta_{St}}}{\partial t} \sin{\alpha_{St}}
\end{equation}

\begin{equation}
\label{eq:local_slope_stokes}
\alpha_{St} = \tan^{-1}{\dfrac{\partial \eta_{St}}{\partial x}}
\end{equation}

\begin{figure}[h]
\label{fig:effective_gravities}
\centering
\includegraphics[width=0.6\textwidth]{figures/fig_effective_gravities.png}
\caption{
  (a) Analytical solutions for the effective gravitational acceleration
  modulation by long waves, as function of long-wave phase, for $\varepsilon_L = 0.2$.
  Black is for linear wave evaluated at $z=0$,
  blue and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and without curvature effects,
  and orange and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and with curvature effects.
  (b) As in (a) but normalized by the modulation of the linear wave evaluated at $z=0$.
}
\end{figure}

\begin{figure}[h]
\label{fig:effective_gravities_mean_max_min}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_effective_gravities_mean_max_min.png}
\caption{
  As Fig. \ref{fig:effective_gravities} but showing the mean (solid),
  maximum (dashed), and minimum (dotted) short-wave gravity modulation as
  function of long-wave steepness $\varepsilon_L$.
}
\end{figure}
